### 2004-05-11 Documentation/part2.tex
*** empty log message ***

% continue here (WW):

@@

 The first eight characters have to be {\tt \%!GLoBES} 
 this serves to protect from parsing MBytes of chunk
 and cluttering the screen with error messages.

Comments
 are entered like this one -- everything between the 
 delimiters is ignored. 
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */
}
\end{quote}
thus the comments in AEDL work like in C

There are so called  named quantities, they all start with {\tt \$} 
and are predefined. They are checked for being within range, \eg\ {\tt 
\$bins} can be only in between 0 and 150. If you give a {\tt float} where
 an {\tt int} is expected the {\tt float} gets converted to an {\tt int} 
in the usual way in C.  We have named scalar quantities
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
as well as named simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\}
}
\end{quote}

Since there are some very special data to parse, some special
provisions have been made so called named environments. This is illustrated 
with the channel definition part
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}

The first part tells the type of environment, here it is {\tt channel}. 
There are the following types 
\begin{quote}
{\tt flux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the type there is a user-defined name beginning with {\tt \#}, in the
above example {\tt \#ch1}, which can be used later on to refer to the 
channel defined in between {\tt <$\ldots$>}. Those names are so called
automatic variables. They start all with {\tt \#} and do not behave as 
usual variables. New names can only be
created within the above construction. Only names which have been created
can be refered to. Using the same name for different types of environments 
does not work.  Each type has its own counter, \eg\ for channel the counter is 
{\tt numofchannels}. The counter keeps track of how many different names 
there are for one type, \ie\ they count the number of channels, 
the number of rules, the number of energy resolution functions asf. 
The value of those automatic variables is the number which can be used 
to refer to them in your C code. You can create a variable for later use by
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now you can refer to {\tt \#ch2} without any trouble and you now could define
a rule using {\tt \#ch2}. Later on of course you still have to specify the
actual channel as described in section~\ref{sec:channel}.

Variables beginning with {\tt @} are named and can be used only within the 
appropriate type of environment, since we have to keep track where 
they belong to. Moreover many of them have a special syntax, like 
{\tt @channel}.

You can have several experiments in one file, just separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope only within a region delimited
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
the counters defined with {\tt cross} and {\tt flux}. They are valid for 
the complete session, since you never must overload the flux and 
cross section information. This ensures  a consistent treatment of various 
experiments in one file - there is no  shadowing of any variables.

Using multiple files to keep things clean and clear is supported by 
 an {\tt include} command. Includes can be nested up 
 to {\tt MAX\_INCLUSION\_DEPTH}, which is set to 10. Error reporting works 
 also for nested includes. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste
\begin{quote}
{\tt include ./file\_1}
\end{quote}

Thus it is possible and advised to use constructs as the following
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
to define a combined analysis of the experiments defined in the files 
{\tt NuFact.gls} and {\tt JHFHK.gls}. Even if you use the named 
variable {\tt \#CC} in both experiments, but the cross section data used 
is different (\eg\ because of different target nuclei), the correct 
cross section data is used for each of the experiments, since {\tt \#CC} 
in {\tt NuFact.gls} has the value $0$ and
in {\tt JHFHK.gls} it has the value $1$, thus no shadowing occurs.

Furthermore you can store numbers in variables
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
And additionally we have some simple calculator functionality like
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available 
{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
Those calculations work everywhere, where
a simple number could stand. However threading over lists does not
 work, \eg\ {\tt sin(\{1,2,3\})} will not work.

One nasty thing is that a line feed character \verb+\n+ is necessary at
 the end of input, alternatively you can put a comment at the end.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties}
\label{sec:source}

@@

The format of the flux files is seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own normalization 
factor as specified by {\tt norm}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. The fluxes
will be rescaled by $1/L^2$, thus the normalization must contain a factor
$L_0^2$, when $L_0$ is the distance from the source for which the flux is
given.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oscillation properties}

\subsection{Baseline \& matter profile}
The purpose of this module is to specify the path the neutrinos
are traveling. The central piece of information is the total length
of the path - the baseline\index{Baseline}
\begin{equation}
L\,\left[\mathrm{km}\right]
\end{equation} 
The baseline is set by 
\begin{quote}
{\tt \$baseline = 3000.0 }
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Detection properties}

\subsection{Cross section}
\label{sec:cross_section}

@@

\subsection{Channels}
\label{sec:channel}

@@

For the calculation of event rates, the first step is to compute the number of
events for each IT in the fiducial mass of the detector for each 
neutrino flavor and energy bin. The second step is to include the detector 
effects coming from
the insufficient knowledge used in the event reconstruction. We combine these
two steps in order to obtain the differential event rate spectrum for each
flavor and IT as seen by the detector, which we call a ``channel''. 
The master formula for the differential event rate for each channel 
is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
\label{eq:master_event}
\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int \int dE\,d\hat{E}\quad
\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
\theta_{12},\theta_{13},
\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\alpha$ is the initial flavor of the neutrino and 
$\beta$ is the final flavour, 
$E$ is the incident neutrino energy, $\Phi_{\alpha} (E)$ is the flux of the 
initial flavor at the
source, $L$ is the baseline length, $N$ is a normalization factor, and 
$\rho$ is the matter density. The interaction term is composed of 
two factors, which are the total cross section 
$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
the interaction type IT, and the energy distribution of the 
secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$ with 
$\hat{E}$ the energy of the secondary particle. The detector properties are 
modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
limited resolution or the cuts in the analysis, and the energy resolution 
function $V_\beta(\hat{E}-E')$ of the secondary particle. Thus, $E'$ is the 
{\em reconstructed} neutrino energy.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rules}
\label{sec:rules}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy resolution function}
\label{sec:energy}

@@

{\tt 
energy(\#name)<\\
\tb @energy =   \{0,2, 0.8634265, 0.0682827,     4e-06\}:\\
\tb\tb \{0,4, 0.1507103, 0.6965592, 0.1507103,   0.00101,     1e-07\}:\\
\tb\tb $\ldots$\\
>
}
\end{quote}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User-defined systematics}
\label{sec:ud_error_dim}.

>>>>

@@

We now give a short introduction to the syntax of AEDL.
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages. 
%
Comments can be used such as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $150$. If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environment in AEDL:
\begin{quote}
{\tt flux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
However, similar to C, one can give a header definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). Within each environment type, there are several 
variables beginning with {\tt @}, which can only be used within the 
appropriate type of environment. In many cases, 
they have a special syntax, such as {\tt @channel}.

If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
the counters defined by {\tt cross} and {\tt flux}. They are valid for 
the complete session, since one does not want too overload the flux and 
cross section information when using more than one experiment. This allows 
a consistent treatment of various experiments in one file.

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include ./file\_1}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
in order to be able to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls} later. 
Even if one uses the
automatic variable {\tt \#CC} in both experiments 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments in the 
application software. In this case, {\tt \#CC} in {\tt NuFact.gls} has 
the value $0$ and in {\tt JHFHK.gls} it has the value $1$, 
which means that no shadowing occurs.

Furthermore, one can define constants such as
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
and use some simple algebraic manipulations such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, such as {\tt sin(\{1,2,3\})} will not work.

Finally, note that a line feed character \verb+\n+ is necessary at
 the end of the input -- alternatively you can put a comment at the end.

% Continue here (WW):

@@

In this chapter, we give a detailed definition of the AEDL features. We also show the underlying mathematical concepts in parallel, where applicable. We do not exactly follow the separation of source, oscillation, and detection properties, since most issues more or less involve the detection. Instead,
we illustrate many of the features of the \GLOBES\ simulation by successively
evolving the underlying statistical picture, and demonstrating how it translates into AEDL .

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties and luminosity}
\label{sec:source}

@@

The format of the flux files is seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own normalization 
factor as specified by {\tt norm}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. The fluxes
will be rescaled by $1/L^2$, thus the normalization must contain a factor
$L_0^2$, when $L_0$ is the distance from the source for which the flux is
given.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Baseline and matter density profile}

The purpose of this module is to specify the path the neutrinos
are traveling. The central piece of information is the total length
of the path - the baseline\index{Baseline}
\begin{equation}
L\,\left[\mathrm{km}\right]
\end{equation} 
The baseline is set by 
\begin{quote}
{\tt \$baseline = 3000.0 }
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cross sections}
\label{sec:cross_section}

@@

\section{Oscillation channels}
\label{sec:channel}

@@

For the calculation of event rates, the first step is to compute the number of
events for each IT in the fiducial mass of the detector for each 
neutrino flavor and energy bin. The second step is to include the detector 
effects coming from
the insufficient knowledge used in the event reconstruction. We combine these
two steps in order to obtain the differential event rate spectrum for each
flavor and IT as seen by the detector, which we call a ``channel''.

The master formula for the differential event rate for each channel 
is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
\label{eq:master_event}
\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int \int dE\,d\hat{E}\quad
\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
\theta_{12},\theta_{13},
\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\alpha$ is the initial flavor of the neutrino and 
$\beta$ is the final flavour, 
$E$ is the incident neutrino energy, $\Phi_{\alpha} (E)$ is the flux of the 
initial flavor at the
source, $L$ is the baseline length, $N$ is a normalization factor, and 
$\rho$ is the matter density. The interaction term is composed of 
two factors, which are the total cross section 
$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
the interaction type IT, and the energy distribution of the 
secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$ with 
$\hat{E}$ the energy of the secondary particle. The detector properties are 
modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
limited resolution or the cuts in the analysis, and the energy resolution 
function $V_\beta(\hat{E}-E')$ of the secondary particle. Thus, $E'$ is the 
{\em reconstructed} neutrino energy.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Energy resolution function}
\label{sec:energy}

@@

{\tt 
energy(\#name)<\\
\tb @energy =   \{0,2, 0.8634265, 0.0682827,     4e-06\}:\\
\tb\tb \{0,4, 0.1507103, 0.6965592, 0.1507103,   0.00101,     1e-07\}:\\
\tb\tb $\ldots$\\
>
}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rules and the treatment of systematics}
\label{sec:rules}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{User-defined systematics}
%\label{sec:ud_error_dim}.
